---
title: "p8105_hw6_rh3064.Rmd"
author: "Rahul Hosalli"
date: "`r Sys.Date()`"
output: github_document
---

### Dependencies

```{r include=FALSE}
library(tidyverse)
library(rvest)
library(httr)
```

## Problem 2

The data is loaded and a city_check function is created to check cities and states in the dataset.

```{r}
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/cbcb572ed1a3c3f10d6dfe07e84464a28986173e/homicide-data.csv"
wphom_df <- read_csv(url)

city_check <- function(df){
  df_sum <- df %>% 
    group_by(state) %>% 
    count(city)
  
  df_sum %>% 
    knitr::kable()
}


city_check(wphom_df)
```

A New city_state variable is created, and WI is correct to be uppercase. A new binary resolved variable is created, and victim_age and victim_race are coerced into the correct type. Cities that do not report victim race are filtered out, along with victims that are neither black nor white. city_check is used to check if the filtering was performed correctly, along with count().

```{r}
wphom_tidy <- wphom_df%>% 
  mutate(
    state = case_when(
      city == "Milwaukee" ~ "WI",
      TRUE ~ state
    ),
    
    city_state = paste(city, state, sep = ", "),
    
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")
  )

wphom_final <- wphom_tidy %>%
  filter(!city_state %in%  c("Dallas, TX", "Phoenix, AZ", 
                         "Kansas City, MO", "Tulsa, AL") &
         victim_race %in% c("Black", "White"))
city_check(wphom_final)
wphom_final %>%
  group_by(victim_race) %>%
  count() %>%
  knitr::kable()
```
